---
title: "Net calcification"
author: "Jill Ashey"
date: "2025-08-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(emmeans)
library(lme4)
library(lmerTest)
library(lsmeans)
```

Read in colony metadata
```{r}
metadata <- read.csv("../data/02-Physiology/Pairs_metadata.csv")
metadata$ColonyID <- as.character(metadata$ColonyID)
```

Read in wax surface area data 
```{r}
surface_area <- read.csv("../output/02-Physiology/SurfaceArea/SurfaceArea_wax_calculations.csv")
surface_area$ColonyID <- as.character(surface_area$ColonyID)
```

Read in data 
```{r}
calc <- read.csv("../data/02-Physiology/Calcification/Calcification_data_field.csv", header = T)
```

Join with colony metadata
```{r}
calc <- calc %>%
  left_join(metadata, by = "ColonyID")
```

Add Timepoint as a column
```{r}
calc <- calc %>%
  mutate(Timepoint = case_when(
    grepl("2023-04", Date) ~ "April_2023",
    grepl("2023-09", Date) ~ "September_2023",
    grepl("2024-01", Date) ~ "January_2024",
    grepl("2024-05", Date) ~ "May_2024",
    grepl("2024-10", Date) ~ "October_2024",
    grepl("2025-01", Date) ~ "January_2025",
    grepl("2025-05", Date) ~ "May_2025",
    grepl("2025-10", Date) ~ "October_2025",
    grepl("2026-01", Date) ~ "January_2026",
    grepl("2026-05", Date) ~ "May_2026",
    TRUE ~ NA_character_))
```

Calculate incubation time 
```{r}
calc <- calc %>%
  mutate(
    start = hm(Time_start),
    end   = hm(Time_end),
    elapsed_mins  = as.numeric(end - start, units = "mins"),
    elapsed_hours = elapsed_mins / 60)
```

Calculate delta alkalinity 
```{r}
calc <- calc %>%
  mutate(Delta_TA_raw = Starting_alk_value - Alk_avg)
```

Calculate blanks 
```{r}
blank_delta_ta <- calc %>%
  filter(ColonyID == "blank" | Notes == "blank") %>%
  group_by(Timepoint) %>%
  summarise(mean_blank_delta_ta = mean(Delta_TA_raw, na.rm = TRUE))# %>%
  #pull(mean_blank_delta_ta)

# Add manual substitutions for missing timepoints
blank_subs <- tibble(
  Timepoint = c("April_2023", "September_2023"),
  mean_blank_delta_ta = c(
    blank_delta_ta$mean_blank_delta_ta[blank_delta_ta$Timepoint == "May_2024"],
    blank_delta_ta$mean_blank_delta_ta[blank_delta_ta$Timepoint == "October_2024"]
  )
)

# Bind the substitutions into the blank dataset
blank_delta_ta <- bind_rows(blank_delta_ta, blank_subs)
```

Some of the timepoints (April and Sept 2023) do not have alk data for the blanks. Instead, use the May and Oct 2024 blanks instead. 

Subtract blank and calculate net calcification
```{r}
# Correct Delta_TA by subtracting blank
calc <- calc %>%
  left_join(blank_delta_ta, by = "Timepoint") 

calc <- calc %>%
  mutate(
    Delta_TA_corr = Delta_TA_raw - mean_blank_delta_ta,
    # Assuming seawater density 1.028 g/mL and volume in mL to L, and no coral weight (thus calc mg/L/hr)
    Net_Calcification = (Delta_TA_corr / 2) * (Sample_Volume_mL / 1000) * 1.028 / elapsed_hours) %>%
  filter(!is.na(Net_Calcification) & !is.nan(Net_Calcification))
```

Read in wax surface area data 
```{r}
surface_area <- read.csv("../output/02-Physiology/SurfaceArea/SurfaceArea_wax_calculations.csv")
surface_area$ColonyID <- as.character(surface_area$ColonyID)

# Subset specific columns
surface_area <- surface_area %>%
  select(ColonyID, Species, Phenotype, surface.area.cm2, Timepoint)

#surface_area <- surface_area %>%
#  select(ColonyID, Species, Phenotype)
```

Join dfs 
```{r}
joined <- calc %>%
  full_join(surface_area, by = c("ColonyID", "Timepoint")) %>%
  mutate(surface.area.cm2 = ifelse(is.na(surface.area.cm2), 10, surface.area.cm2)) %>%
  filter(ColonyID != "blank" & !is.na(ColonyID)) %>%
  rename(Phenotype = Phenotype.x, Species = Species.x) %>%
  select(Date, Timepoint, Light, ColonyID, Species, Phenotype, Net_Calcification, surface.area.cm2)

#joined <- calc %>%
#  left_join(surface_area, by = c("ColonyID", "Timepoint"))
```

Normalize by SA
```{r}
joined <- joined %>%
  mutate(Net_Calcification_cm2 = Net_Calcification / surface.area.cm2)
```

Save as csv 
```{r}
write.csv(joined, "../output/02-Physiology/Calcification/Calcification_data_normalized.csv")
```

Calculate mean, sem, and sd 
```{r}
joined_summary <- joined %>%
  group_by(Timepoint, Phenotype, Species, Light) %>%
  summarise(
    mean_calc = mean(Net_Calcification_cm2, na.rm = TRUE),
    sd_calc = sd(Net_Calcification_cm2, na.rm = T),
    se_calc = sd(Net_Calcification_cm2, na.rm = T) / sqrt(length(Net_Calcification_cm2)),
    .groups = "drop") %>%
  drop_na()# %>%
  #filter(Light == "Light")
  #filter(!Timepoint == "Jan25") # remove Jan25 for now 
```

Plot! 
```{r}
# Set levels 
joined_summary$Timepoint <- factor(joined_summary$Timepoint, levels = c("April_2023",
                                                                        "September_2023",
                                                                        "January_2024", 
                                                                        "May_2024", 
                                                                        "October_2024", 
                                                                        "January_2025", 
                                                                        "May_2025",
                                                                        "October_2025",
                                                                        "January_2026",
                                                                        "May_2026"))

ggplot(joined_summary, aes(x = Timepoint, y = mean_calc, color = Phenotype, group = Phenotype)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_calc - se_calc, ymax = mean_calc + se_calc, color = Phenotype),
                width = 0.3, linewidth = 0.8) +  # color matches points/lines
  #facet_wrap(~ Species + Light) +
  facet_grid(rows = vars(Species), cols = vars(Light)) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")
  ) +
  theme_classic(base_size = 16) +  # bigger base size
  labs(
    x = "",
    y = "Net calcification (umol/hr/cm2)",
    color = "Phenotype"
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "italic", size = 16),
    #legend.position = c(0.85, 0.83),  # adjust x, y for placement
    legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 13, color = "black"),
    axis.text.y = element_text(size = 13, color = "black"),
    axis.title.y = element_text(size = 16)
  )
```

Plot light only 
```{r}
joined_summary_light <- joined_summary %>%
  filter(Light == "Light")

joined_summary_light$Timepoint <- gsub("_", " ", joined_summary_light$Timepoint)

# Set levels 
joined_summary_light$Timepoint <- factor(joined_summary_light$Timepoint, levels = c("April 2023",
                                                                        "September 2023",
                                                                        "January 2024", 
                                                                        "May 2024", 
                                                                        "October 2024", 
                                                                        "January 2025", 
                                                                        "May 2025",
                                                                        "October 2025",
                                                                        "January 2026",
                                                                        "May 2026"))

y_label <- expression(bold(paste("Net calcification (", mu, "mol ", cm^-2, " ", hr^-1, ")")))
light_calc <- ggplot(joined_summary_light, aes(x = Timepoint, y = mean_calc, color = Phenotype, group = Phenotype)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +
  geom_line(linewidth = 2.2) +
  geom_point(size = 7) +
  geom_errorbar(aes(ymin = mean_calc - se_calc, ymax = mean_calc + se_calc, color = Phenotype),
                width = 0.4, linewidth = 1.8) +  # color matches points/lines
  #facet_wrap(~ Species + Light) +
  facet_grid(cols = vars(Species)) +
  scale_y_continuous(breaks = seq(-2, 10, by = 0.2)) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")
  ) +
  theme_classic() +  
  labs(
    x = "",
    y = y_label,
    color = "Phenotype"
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "bold.italic", size = 24),
    #legend.position = c(0.85, 0.83),  # adjust x, y for placement
    legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 18, face = "bold"),
    legend.text = element_text(size = 17),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 20, color = "black"),
    axis.text.y = element_text(size = 24, color = "black"),
    axis.title.y = element_text(size = 26, face = "bold")
); light_calc

ggsave("../output/02-Physiology/Calcification/Calcification_light_KBay.png", 
       plot = light_calc, width = 20, height = 8, units = "in", dpi = 300)
ggsave("../output/02-Physiology/Calcification/Calcification_light_KBay.pdf", 
       plot = light_calc, width = 20, height = 8, units = "in", device = "pdf", dpi = 300)
```

Plot separately by spp 
```{r}
### Mcap 
joined_summary_light_mcap <- joined_summary_light %>%
  filter(Species == "Montipora capitata")

# Set levels 
joined_summary_light_mcap$Timepoint <- factor(joined_summary_light_mcap$Timepoint, levels = c("April 2023",
                                                                        "September 2023",
                                                                        "January 2024", 
                                                                        "May 2024", 
                                                                        "October 2024", 
                                                                        "January 2025", 
                                                                        "May 2025",
                                                                        "October 2025",
                                                                        "January 2026",
                                                                        "May 2026"))

y_label <- expression(bold(paste("Net calcification (", mu, "mol ", cm^-2, " ", hr^-1, ")")))
light_calc_mcap <- ggplot(joined_summary_light_mcap, aes(x = Timepoint, y = mean_calc, color = Phenotype, group = Phenotype)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +
  geom_line(linewidth = 2.2) +
  geom_point(size = 7) +
  geom_errorbar(aes(ymin = mean_calc - se_calc, ymax = mean_calc + se_calc, color = Phenotype),
                width = 0.4, linewidth = 1.8) +  # color matches points/lines
  #facet_wrap(~ Species + Light) +
  #facet_grid(cols = vars(Species)) +
  scale_y_continuous(breaks = seq(0, 0.5, by = 0.1)) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")
  ) +
  theme_classic() +  
  labs(
    x = "",
    y = y_label,
    color = "Phenotype"
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "bold.italic", size = 24),
    #legend.position = c(0.85, 0.83),  # adjust x, y for placement
    legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 18, face = "bold"),
    legend.text = element_text(size = 17),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 20, color = "black"),
    axis.text.y = element_text(size = 24, color = "black"),
    axis.title.y = element_text(size = 26, face = "bold")
); light_calc_mcap

ggsave("../output/02-Physiology/Calcification/Calcification_light_KBay_Mcap.png", 
       plot = light_calc_mcap, width = 15, height = 10, units = "in", dpi = 300)
ggsave("../output/02-Physiology/Calcification/Calcification_light_KBay_Mcap.pdf", 
       plot = light_calc_mcap, width = 15, height = 10, units = "in", device = "pdf", dpi = 300)

### Pcomp 
joined_summary_light_pcomp <- joined_summary_light %>%
  filter(Species == "Porites compressa")

# Set levels 
joined_summary_light_pcomp$Timepoint <- factor(joined_summary_light_pcomp$Timepoint, levels = c("April 2023",
                                                                        "September 2023",
                                                                        "January 2024", 
                                                                        "May 2024", 
                                                                        "October 2024", 
                                                                        "January 2025", 
                                                                        "May 2025",
                                                                        "October 2025",
                                                                        "January 2026",
                                                                        "May 2026"))

y_label <- expression(bold(paste("Net calcification (", mu, "mol ", cm^-2, " ", hr^-1, ")")))
light_calc_pcomp <- ggplot(joined_summary_light_pcomp, aes(x = Timepoint, y = mean_calc, color = Phenotype, group = Phenotype)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +
  geom_line(linewidth = 2.2) +
  geom_point(size = 7) +
  geom_errorbar(aes(ymin = mean_calc - se_calc, ymax = mean_calc + se_calc, color = Phenotype),
                width = 0.4, linewidth = 1.8) +  # color matches points/lines
  #facet_wrap(~ Species + Light) +
  #facet_grid(cols = vars(Species)) +
  scale_y_continuous(breaks = seq(-2, 10, by = 0.2)) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")
  ) +
  theme_classic() +  
  labs(
    x = "",
    y = y_label,
    color = "Phenotype"
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "bold.italic", size = 24),
    #legend.position = c(0.85, 0.83),  # adjust x, y for placement
    legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 18, face = "bold"),
    legend.text = element_text(size = 17),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 20, color = "black"),
    axis.text.y = element_text(size = 24, color = "black"),
    axis.title.y = element_text(size = 26, face = "bold")
); light_calc_pcomp

ggsave("../output/02-Physiology/Calcification/Calcification_light_KBay_Pcomp.png", 
       plot = light_calc_pcomp, width = 15, height = 10, units = "in", dpi = 300)
ggsave("../output/02-Physiology/Calcification/Calcification_light_KBay_Pcomp.pdf", 
       plot = light_calc_pcomp, width = 15, height = 10, units = "in", device = "pdf", dpi = 300)
```

Average across season 
```{r}
joined_season_summary <- joined %>%
  mutate(Season = case_when(
    Timepoint == "April_2023" ~ "Spring",
    Timepoint == "September_2023" ~ "Fall",
    Timepoint == "January_2024" ~ "Winter",
    Timepoint == "May_2024" ~ "Spring", 
    Timepoint == "October_2024" ~ "Fall",
    Timepoint == "January_2025" ~ "Winter", 
    Timepoint == "May_2025" ~ "Spring",
    Timepoint == "October_2025" ~ "Fall", 
    Timepoint == "January_2026" ~ "Winter",
    Timepoint == "May_2026" ~ "Spring")) %>%
  group_by(Season, Phenotype, Species, Light) %>%
  summarise(
    mean_calc = mean(Net_Calcification_cm2, na.rm = TRUE),
    sd_calc = sd(Net_Calcification_cm2, na.rm = T),
    se_calc = sd(Net_Calcification_cm2, na.rm = T) / sqrt(length(Net_Calcification_cm2)),
    .groups = "drop") %>%
  drop_na()
```

Plot across season by spp 
```{r}
### Mcap 
joined_season_light_summary_mcap <- joined_season_summary %>%
  filter(Species == "Montipora capitata") %>%
  filter(Light == "Light")

# Set levels 
joined_season_light_summary_mcap$Season <- factor(joined_season_light_summary_mcap$Season, levels = c("Winter", "Spring", "Fall"))

y_label <- expression(bold(paste("Net calcification (", mu, "mol ", cm^-2, " ", hr^-1, ")")))
calc_season_plot_mcap <- ggplot(joined_season_light_summary_mcap, aes(x = Season, y = mean_calc, color = Phenotype, group = Phenotype)) +
  #geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +
  geom_line(linewidth = 2.2) +
  geom_point(size = 7) +
  geom_errorbar(aes(ymin = mean_calc - se_calc, ymax = mean_calc + se_calc, color = Phenotype),
                width = 0.4, linewidth = 1.8) +  # color matches points/lines
  #facet_wrap(~ Species + Light) +
  #facet_grid(cols = vars(Species)) +
  #scale_y_continuous(breaks = seq(-0.3, -1.3, by = -0.2)) +
  scale_y_continuous() +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")
  ) +
  theme_classic() +  
  labs(
    x = "",
    y = y_label,
    color = "Phenotype"
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "bold.italic", size = 24),
    #legend.position = c(0.85, 0.83),  # adjust x, y for placement
    legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 18, face = "bold"),
    legend.text = element_text(size = 17),
    axis.text.x = element_text(hjust = 0.5, vjust = 0, size = 20, color = "black"),
    axis.text.y = element_text(size = 22, color = "black"),
    axis.title.y = element_text(size = 24, face = "bold")
); calc_season_plot_mcap

ggsave("../output/02-Physiology/Calcification/Calcification_light_seasonal_KBay_Mcap.png", 
       plot = calc_season_plot_mcap, width = 12, height = 8, units = "in", dpi = 300)
ggsave("../output/02-Physiology/Calcification/Calcification_light_seasonal_KBay_Mcap.pdf", 
       plot = calc_season_plot_mcap, width = 12, height = 8, units = "in", device = "pdf", dpi = 300)

### Pcomp 
joined_season_light_summary_pcomp <- joined_season_summary %>%
  filter(Species == "Porites compressa") %>%
  filter(Light == "Light")

# Set levels 
joined_season_light_summary_pcomp$Season <- factor(joined_season_light_summary_pcomp$Season, levels = c("Winter", "Spring", "Fall"))

y_label <- expression(bold(paste("Net calcification (", mu, "mol ", cm^-2, " ", hr^-1, ")")))
calc_season_plot_pcomp <- ggplot(joined_season_light_summary_pcomp, aes(x = Season, y = mean_calc, color = Phenotype, group = Phenotype)) +
  #geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +
  geom_line(linewidth = 2.2) +
  geom_point(size = 7) +
  geom_errorbar(aes(ymin = mean_calc - se_calc, ymax = mean_calc + se_calc, color = Phenotype),
                width = 0.4, linewidth = 1.8) +  # color matches points/lines
  #facet_wrap(~ Species + Light) +
  #facet_grid(cols = vars(Species)) +
  #scale_y_continuous(breaks = seq(-0.3, -1.3, by = -0.2)) +
  scale_y_continuous() +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")
  ) +
  theme_classic() +  
  labs(
    x = "",
    y = y_label,
    color = "Phenotype"
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "bold.italic", size = 24),
    #legend.position = c(0.85, 0.83),  # adjust x, y for placement
    legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 18, face = "bold"),
    legend.text = element_text(size = 17),
    axis.text.x = element_text(hjust = 0.5, vjust = 0, size = 20, color = "black"),
    axis.text.y = element_text(size = 22, color = "black"),
    axis.title.y = element_text(size = 24, face = "bold")
); calc_season_plot_pcomp

ggsave("../output/02-Physiology/Calcification/Calcification_light_seasonal_KBay_Pcomp.png", 
       plot = calc_season_plot_pcomp, width = 12, height = 8, units = "in", dpi = 300)
ggsave("../output/02-Physiology/Calcification/Calcification_light_seasonal_KBay_Pcomp.pdf", 
       plot = calc_season_plot_pcomp, width = 12, height = 8, units = "in", device = "pdf", dpi = 300)
```






















Plot dark only 
```{r}
joined_summary_dark <- joined_summary %>%
  filter(Light == "Dark")

joined_summary_dark$Timepoint <- gsub("_", " ", joined_summary_dark$Timepoint)

# Set levels 
joined_summary_dark$Timepoint <- factor(joined_summary_dark$Timepoint, levels = c("April 2023",
                                                                        "September 2023",
                                                                        "January 2024", 
                                                                        "May 2024", 
                                                                        "October 2024", 
                                                                        "January 2025", 
                                                                        "May 2025",
                                                                        "October 2025",
                                                                        "January 2026",
                                                                        "May 2026"))

y_label <- expression(bold(paste("Net calcification (", mu, "mol ", cm^-2, " ", hr^-1, ")")))
dark_calc <- ggplot(joined_summary_dark, aes(x = Timepoint, y = mean_calc, color = Phenotype, group = Phenotype)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +
  geom_line(linewidth = 2.2) +
  geom_point(size = 7) +
  geom_errorbar(aes(ymin = mean_calc - se_calc, ymax = mean_calc + se_calc, color = Phenotype),
                width = 0.4, linewidth = 1.8) +  # color matches points/lines
  #facet_wrap(~ Species + Light) +
  facet_grid(cols = vars(Species)) +
  scale_y_continuous(breaks = seq(-2, 10, by = 0.2)) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")
  ) +
  theme_classic() +  
  labs(
    x = "",
    y = y_label,
    color = "Phenotype"
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "bold.italic", size = 24),
    #legend.position = c(0.85, 0.83),  # adjust x, y for placement
    legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 18, face = "bold"),
    legend.text = element_text(size = 17),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 20, color = "black"),
    axis.text.y = element_text(size = 24, color = "black"),
    axis.title.y = element_text(size = 26, face = "bold")
); dark_calc

ggsave("../output/02-Physiology/Calcification/Calcification_dark_KBay.png", 
       plot = dark_calc, width = 20, height = 8, units = "in", dpi = 300)
ggsave("../output/02-Physiology/Calcification/Calcification_dark_KBay.pdf", 
       plot = dark_calc, width = 20, height = 8, units = "in", device = "pdf", dpi = 300)
```

Plot separately by spp 
```{r}
### Mcap 
joined_summary_dark_mcap <- joined_summary_dark %>%
  filter(Species == "Montipora capitata")

# Set levels 
joined_summary_dark_mcap$Timepoint <- factor(joined_summary_dark_mcap$Timepoint, levels = c("April 2023",
                                                                        "September 2023",
                                                                        "January 2024", 
                                                                        "May 2024", 
                                                                        "October 2024", 
                                                                        "January 2025", 
                                                                        "May 2025",
                                                                        "October 2025",
                                                                        "January 2026",
                                                                        "May 2026"))

y_label <- expression(bold(paste("Net calcification (", mu, "mol ", cm^-2, " ", hr^-1, ")")))
dark_calc_mcap <- ggplot(joined_summary_dark_mcap, aes(x = Timepoint, y = mean_calc, color = Phenotype, group = Phenotype)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +
  geom_line(linewidth = 2.2) +
  geom_point(size = 7) +
  geom_errorbar(aes(ymin = mean_calc - se_calc, ymax = mean_calc + se_calc, color = Phenotype),
                width = 0.4, linewidth = 1.8) +  # color matches points/lines
  #facet_wrap(~ Species + Light) +
  #facet_grid(cols = vars(Species)) +
  scale_y_continuous(breaks = seq(0, 0.5, by = 0.1)) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")
  ) +
  theme_classic() +  
  labs(
    x = "",
    y = y_label,
    color = "Phenotype"
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "bold.italic", size = 24),
    #legend.position = c(0.85, 0.83),  # adjust x, y for placement
    legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 18, face = "bold"),
    legend.text = element_text(size = 17),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 20, color = "black"),
    axis.text.y = element_text(size = 24, color = "black"),
    axis.title.y = element_text(size = 26, face = "bold")
); dark_calc_mcap

ggsave("../output/02-Physiology/Calcification/Calcification_dark_KBay_Mcap.png", 
       plot = dark_calc_mcap, width = 15, height = 10, units = "in", dpi = 300)
ggsave("../output/02-Physiology/Calcification/Calcification_dark_KBay_Mcap.pdf", 
       plot = dark_calc_mcap, width = 15, height = 10, units = "in", device = "pdf", dpi = 300)

### Pcomp 
joined_summary_dark_pcomp <- joined_summary_dark %>%
  filter(Species == "Porites compressa")

# Set levels 
joined_summary_dark_pcomp$Timepoint <- factor(joined_summary_dark_pcomp$Timepoint, levels = c("April 2023",
                                                                        "September 2023",
                                                                        "January 2024", 
                                                                        "May 2024", 
                                                                        "October 2024", 
                                                                        "January 2025", 
                                                                        "May 2025",
                                                                        "October 2025",
                                                                        "January 2026",
                                                                        "May 2026"))

y_label <- expression(bold(paste("Net calcification (", mu, "mol ", cm^-2, " ", hr^-1, ")")))
dark_calc_pcomp <- ggplot(joined_summary_dark_pcomp, aes(x = Timepoint, y = mean_calc, color = Phenotype, group = Phenotype)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +
  geom_line(linewidth = 2.2) +
  geom_point(size = 7) +
  geom_errorbar(aes(ymin = mean_calc - se_calc, ymax = mean_calc + se_calc, color = Phenotype),
                width = 0.4, linewidth = 1.8) +  # color matches points/lines
  #facet_wrap(~ Species + Light) +
  #facet_grid(cols = vars(Species)) +
  scale_y_continuous(breaks = seq(-2, 10, by = 0.2)) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")
  ) +
  theme_classic() +  
  labs(
    x = "",
    y = y_label,
    color = "Phenotype"
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "bold.italic", size = 24),
    #legend.position = c(0.85, 0.83),  # adjust x, y for placement
    legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 18, face = "bold"),
    legend.text = element_text(size = 17),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 20, color = "black"),
    axis.text.y = element_text(size = 24, color = "black"),
    axis.title.y = element_text(size = 26, face = "bold")
); dark_calc_pcomp

ggsave("../output/02-Physiology/Calcification/Calcification_dark_KBay_Pcomp.png", 
       plot = dark_calc_pcomp, width = 15, height = 10, units = "in", dpi = 300)
ggsave("../output/02-Physiology/Calcification/Calcification_dark_KBay_Pcomp.pdf", 
       plot = dark_calc_pcomp, width = 15, height = 10, units = "in", device = "pdf", dpi = 300)
```













## Stats! 

### Light

Subset by light 
```{r}
joined_light <- joined %>%
  filter(Light == "Light") %>%
  filter(!Timepoint == "Jan25") # remove Jan25 for now 
```

Look at normality 
```{r}
hist(joined_light$Net_Calcification_cm2)

# Q-Q Plot
qqnorm(joined_light$Net_Calcification_cm2)
qqline(joined_light$Net_Calcification_cm2, col = "red")

# Shapiro-Wilk Test
shapiro.test(joined_light$Net_Calcification_cm2)
```

Data is not normal, perform data transformation
```{r}
# Log10
joined_light$calc_log <- log10(joined_light$Net_Calcification_cm2 +1) # add 1 to avoid log10(0)
hist(joined_light$calc_log)
qqnorm(joined_light$calc_log)
qqline(joined_light$calc_log, col = "red")
shapiro.test(joined_light$calc_log)

# Sqrt
joined_light$calc_sqrt <- sqrt(joined_light$Net_Calcification_cm2)
hist(joined_light$calc_sqrt)
qqnorm(joined_light$calc_sqrt)
qqline(joined_light$calc_sqrt, col = "red")
shapiro.test(joined_light$calc_sqrt)

# Inverse 
joined_light$calc_inv <- 1 / (joined_light$Net_Calcification_cm2 +1) # add 1 to avoid division by zero
hist(joined_light$calc_inv)
qqnorm(joined_light$calc_inv)
qqline(joined_light$calc_inv, col = "red")
shapiro.test(joined_light$calc_inv)
```

Sqrt and inv produced normality. Proceed with sqrt. 
```{r}
mod_calc <- lmer(calc_sqrt ~ Species * Phenotype * Timepoint + (1 | Colony_ID),
                 data = joined_light,
                 REML = FALSE)
car::Anova(mod_calc, type = 3)
```

Look at posthoc comparisons 
```{r}
em_spp <- emmeans(mod_calc, pairwise ~ Species, adjust = "tukey")
summary(em_spp)

em_time <- emmeans(mod_calc, pairwise ~ Timepoint, adjust = "tukey")
summary(em_time)
```

Calculate mean, sem, and sd for light only
```{r}
joined_light_summary <- joined_light %>%
  group_by(Timepoint, Phenotype, Species) %>%
  summarise(
    mean_calc = mean(Net_Calcification_cm2, na.rm = TRUE),
    sd_calc = sd(Net_Calcification_cm2, na.rm = T),
    se_calc = sd(Net_Calcification_cm2, na.rm = T) / sqrt(length(Net_Calcification_cm2)),
    .groups = "drop") %>%
  drop_na() %>%
  filter(!Timepoint == "Jan25") # remove Jan25 for now 
```

Plot! 
```{r}
# Set levels 
joined_light_summary$Timepoint <- factor(joined_light_summary$Timepoint, levels = c("Apr23",
                                                                        "Sept23",
                                                                        "Jan24", 
                                                                        "May24", 
                                                                        "Oct24", 
                                                                        #"Jan25", 
                                                                        "May25",
                                                                        "Oct25"))

ggplot(joined_light_summary, aes(x = Timepoint, y = mean_calc, color = Phenotype, group = Phenotype)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_calc - se_calc, ymax = mean_calc + se_calc, color = Phenotype),
                width = 0.3, linewidth = 0.8) +  # color matches points/lines
  facet_wrap(~ Species) +
  #facet_grid(rows = vars(Species), cols = vars(Light)) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")
  ) +
  theme_classic(base_size = 16) +  # bigger base size
  labs(
    x = "",
    y = "Net calcification (Light)",
    color = "Phenotype"
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "italic", size = 16),
    legend.position = c(0.15, 0.83),  # adjust x, y for placement
    #legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 13, color = "black"),
    axis.text.y = element_text(size = 13, color = "black"),
    axis.title.y = element_text(size = 16)
  )
```

### Dark

Subset by Dark 
```{r}
joined_dark <- joined %>%
  filter(Light == "Dark") %>%
  filter(!Timepoint == "Jan25") # remove Jan25 for now 
```

Look at normality 
```{r}
hist(joined_dark$Net_Calcification_cm2)

# Q-Q Plot
qqnorm(joined_dark$Net_Calcification_cm2)
qqline(joined_dark$Net_Calcification_cm2, col = "red")

# Shapiro-Wilk Test
shapiro.test(joined_dark$Net_Calcification_cm2)
```

Data is not normal, perform data transformation
```{r}
# Log10
joined_dark$calc_log <- log10(joined_dark$Net_Calcification_cm2 +1) # add 1 to avoid log10(0)
hist(joined_dark$calc_log)
qqnorm(joined_dark$calc_log)
qqline(joined_dark$calc_log, col = "red")
shapiro.test(joined_dark$calc_log)

# Sqrt
joined_dark$calc_sqrt <- sqrt(joined_dark$Net_Calcification_cm2)
hist(joined_dark$calc_sqrt)
qqnorm(joined_dark$calc_sqrt)
qqline(joined_dark$calc_sqrt, col = "red")
shapiro.test(joined_dark$calc_sqrt)

# Inverse 
joined_dark$calc_inv <- 1 / (joined_dark$Net_Calcification_cm2 +1) # add 1 to avoid division by zero
hist(joined_dark$calc_inv)
qqnorm(joined_dark$calc_inv)
qqline(joined_dark$calc_inv, col = "red")
shapiro.test(joined_dark$calc_inv)
```

Sqrt and inv produced normality. Proceed with sqrt. 
```{r}
mod_calc_dark <- lmer(calc_sqrt ~ Species * Phenotype * Timepoint + (1 | Colony_ID),
                 data = joined_dark,
                 REML = FALSE)
car::Anova(mod_calc_dark, type = 3)
```

Species x Phenotype significant. 

Look at post-hoc results 
```{r}
em_sp_ph <- emmeans(mod_calc_dark, ~ Species * Phenotype | Timepoint)
contrasts_sp_ph <- contrast(em_sp_ph, method = "pairwise", adjust = "tukey")
summary(contrasts_sp_ph)
```

Calculate mean, sem, and sd for dark only
```{r}
joined_dark_summary <- joined_dark %>%
  group_by(Timepoint, Phenotype, Species) %>%
  summarise(
    mean_calc = mean(Net_Calcification_cm2, na.rm = TRUE),
    sd_calc = sd(Net_Calcification_cm2, na.rm = T),
    se_calc = sd(Net_Calcification_cm2, na.rm = T) / sqrt(length(Net_Calcification_cm2)),
    .groups = "drop") %>%
  drop_na() %>%
  filter(!Timepoint == "Jan25") # remove Jan25 for now 
```

Plot! 
```{r}
# Set levels 
joined_dark_summary$Timepoint <- factor(joined_dark_summary$Timepoint, levels = c("Apr23",
                                                                        "Sept23",
                                                                        "Jan24", 
                                                                        "May24", 
                                                                        "Oct24", 
                                                                        #"Jan25", 
                                                                        "May25",
                                                                        "Oct25"))

ggplot(joined_dark_summary, aes(x = Timepoint, y = mean_calc, color = Phenotype, group = Phenotype)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_calc - se_calc, ymax = mean_calc + se_calc, color = Phenotype),
                width = 0.3, linewidth = 0.8) +  # color matches points/lines
  facet_wrap(~ Species) +
  #facet_grid(rows = vars(Species), cols = vars(Light)) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")
  ) +
  theme_classic(base_size = 16) +  # bigger base size
  labs(
    x = "",
    y = "Net calcification (Dark)",
    color = "Phenotype"
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "italic", size = 16),
    legend.position = c(0.85, 0.33),  # adjust x, y for placement
    #legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 13, color = "black"),
    axis.text.y = element_text(size = 13, color = "black"),
    axis.title.y = element_text(size = 16)
  )
```


