---
title: "Physiology"
author: "Jill Ashey"
date: "2025-11-24"
output: html_document
---

Script for all physiology (sym counts, chlorophyll, protein, MAAs, and AFDW) analyses. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(purrr)
library(readr)
library(stringr)
library(emmeans)
```

## Prep for analyses

Run this before running any ANOVAs
```{r}
options(contrasts = c("contr.sum", "contr.poly"))
```

Read in colony metadata
```{r}
metadata <- read.csv("../data/02-Physiology/Pairs_metadata.csv")
metadata$ColonyID <- as.character(metadata$ColonyID)
```

Read in wax surface area data 
```{r}
surface_area <- read.csv("../output/02-Physiology/SurfaceArea/SurfaceArea_wax_calculations.csv")
surface_area$ColonyID <- as.character(surface_area$ColonyID)
```

## Fv/Fm

### Process data 

Read in PAM data
```{r}
pam <- read.csv("../data/02-Physiology/PAM/PAM.csv")
pam$ColonyID <- as.character(pam$ColonyID)
```

Join with colony metadata
```{r}
pam <- pam %>%
  left_join(metadata, by = "ColonyID")
```

Plot data
```{r}
pam <- pam %>%
  filter(!is.na(Fv_Fm)) %>%
  mutate(Timepoint = factor(Timepoint, levels = c(
    "April_2023", 
    "September_2023", 
    "January_2024", 
    "May_2024",
    "October_2024", 
    "January_2025",
    "May_2025",
    "October_2025",
    "January_2026")))

ggplot(pam, aes(x = Timepoint, y = Fv_Fm, fill = Phenotype, color = Phenotype, shape = PAM_type)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.6, position = position_dodge(0.8)) +
  geom_point(aes(color = Phenotype), position = position_dodge(0.8), alpha = 0.6, size = 1.5) +
  facet_wrap(~Species, scales = "free_y") +
  scale_color_manual(values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  scale_fill_manual(values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  theme_bw() +
  labs(
    x = "Timepoint",
    y = "Fv/Fm",
    fill = "Phenotype",
    color = "Phenotype"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold", size = 10),
    legend.position = "right"
  )
```

### Stats

#### Mcap

Subset by Mcap only 
```{r}
pam_mcap <- pam %>%
  filter(Species == "Montipora capitata")
```

Check for normality
```{r}
hist(pam_mcap$Fv_Fm)

# Q-Q Plot
qqnorm(pam_mcap$Fv_Fm)
qqline(pam_mcap$Fv_Fm, col = "red")

# Shapiro-Wilk Test
shapiro.test(pam_mcap$Fv_Fm)
```

Normal enough.

Run linear mixed effects model 
```{r}
pam.mcap.model <- lme4::lmer(Fv_Fm ~ Timepoint * Phenotype + (1|ColonyID), data = pam_mcap)
car::Anova(pam.mcap.model, type = 3)

# Model normality 
par(mfrow = c(1, 2)) 
plot(pam.mcap.model, which = 1) 
lattice::qqmath(pam.mcap.model)

# Post hoc tests 
emmeans(pam.mcap.model, pairwise ~ Timepoint)

# Check plot to make sure results make sense
ggplot(pam_mcap, aes(x = Timepoint, y = Fv_Fm, color = Phenotype, group = Phenotype, shape = PAM_type)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  facet_wrap(~Species) +
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

#### Pcomp

Subset by Pcomp only 
```{r}
pam_pcomp <- pam %>%
  filter(Species == "Porites compressa")
```

Check for normality
```{r}
hist(pam_pcomp$Fv_Fm)

# Q-Q Plot
qqnorm(pam_pcomp$Fv_Fm)
qqline(pam_pcomp$Fv_Fm, col = "red")

# Shapiro-Wilk Test
shapiro.test(pam_pcomp$Fv_Fm)
```

Normal.

Run linear mixed effects model 
```{r}
pam.pcomp.model <- lme4::lmer(Fv_Fm ~ Timepoint * Phenotype + (1|ColonyID), data = pam_pcomp)
car::Anova(pam.pcomp.model, type = 3)

# Model normality 
par(mfrow = c(1, 2)) 
plot(pam.pcomp.model, which = 1) 
lattice::qqmath(pam.pcomp.model)

# Post hoc tests 
emmeans(pam.pcomp.model, pairwise ~ Timepoint)
emmeans(pam.pcomp.model, pairwise ~ Phenotype)
emmeans(pam.pcomp.model, pairwise ~ Timepoint | Phenotype)
emmeans(pam.pcomp.model, pairwise ~ Phenotype | Timepoint)

# Check plot to make sure results make sense
ggplot(pam_pcomp, aes(x = Timepoint, y = Fv_Fm, color = Phenotype, group = Phenotype, shape = PAM_type)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  facet_wrap(~Species) +
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

## Sym counts 

Read in metadata 
```{r}
meta_sym <- read.csv("../data/02-Physiology/Symbiont_density/sym_count_metadata.csv")

# Remove extra spaces 
meta_sym <- meta_sym %>%
  mutate(
    Run = str_replace_all(Run, "\\s+", ""),  # "Run 1" -> "Run1"
    Well = str_trim(Well),
    Date = as.character(Date),
    ColonyID = as.character(ColonyID))
```

Read in raw data from guava
```{r}
files <- list.files(path = "../data/02-Physiology/Symbiont_density/raw_guava", pattern = "*.csv", full.names = TRUE)

read_with_metadata <- function(f) {
  df <- read.csv(f, skip = 4, header = TRUE, stringsAsFactors = FALSE)
  
  # extract filename without path and extension
  filename <- basename(f)
  filename <- str_remove(filename, "\\.csv$")

  # split filename at underscore, e.g., "20230805_Run1" -> c("20230805", "Run1")
  parts <- str_split(filename, "_", simplify = TRUE)

  # add columns for Date and Run parsed from filename
  df %>%
    mutate(
      Date = parts[1],
      Run = parts[2],
      # ensure consistent types and clean whitespace
      Well = str_trim(Well),
      Sample.ID = as.character(Sample.ID)
    )
}

all_dfs <- lapply(files, read_with_metadata)

combined_df <- bind_rows(all_dfs)

combined_df <- combined_df %>%
  mutate(
    Run = str_replace_all(Run, "\\s+", ""),
    Well = str_trim(Well),
    Date = as.character(Date)  # ensure same type
  )
```

Merge raw data and guava metadata. Merge w/ colony metadata as well
```{r}
sym_counts <- left_join(meta_sym, combined_df, by = c("Date", "Run", "Well"))
```

Remove samples marked "Do not use"
```{r}
sym_counts <- sym_counts %>%
  filter(is.na(Use) | Use == "")
```

Calculate mean counts per sample and number of cells per sample 
```{r}
sym_counts_avg <- sym_counts %>%
  left_join(metadata, by = "ColonyID") %>%
  group_by(Date, Run, Dilution, Timepoint, Pellet_Resuspend_Vol_mL, ColonyID) %>%
  summarise(mean_cells_mL = mean(CHLOROPHYLL.POS.Concentration.Concentration.for.CHLOROPHYLL.POS...cells.mL., na.rm = T))%>%
  mutate(cells = Dilution * Pellet_Resuspend_Vol_mL * mean_cells_mL,
         ColonyID = as.character(ColonyID)) %>%
  ungroup()
```


CHECK CALCULATIONS AND ADD manual counts


May 2025
```{r}
sym_counts_avg_may25 <- sym_counts_avg %>%
  filter(Timepoint == "May_2025") %>%
    mutate(cells_cm2 = cells/10,
          cells_cm2_M = cells_cm2 / 1e6)


```





Merge with surface area and calculate cells/cm2
```{r}
sym_counts_avg <- sym_counts_avg %>%
  left_join(surface_area, by = c("ColonyID", "Timepoint")) 

sym_counts_avg <- sym_counts_avg %>%
  mutate(cells_cm2 = cells/surface.area.cm2,
          cells_cm2_M = cells_cm2 / 1e6) %>%
  select(ColonyID, Timepoint, Species, Phenotype, Dilution, cells_cm2, cells_cm2_M)
```

Save as csv 
```{r}
write.csv(sym_counts_avg, "../output/Physiology/sym_density.csv")
```

Calculate mean, sem, and sd 
```{r}
sym_summary <- sym_counts_avg %>%
  group_by(Timepoint, Phenotype, Species) %>%
  summarise(
    mean_sym = mean(cells_cm2_M, na.rm = TRUE),
    sd_sym = sd(cells_cm2_M, na.rm = T),
    se_sym = sd(cells_cm2_M, na.rm = T) / sqrt(length(cells_cm2_M)),
    .groups = "drop") %>%
  drop_na() 
```

Plot! 
```{r}
# Set levels 
sym_summary$Timepoint <- factor(sym_summary$Timepoint, levels = c("April_2023",
                                                                        "September_2023",
                                                                        "January_2024", 
                                                                        "May_2024",
                                                                        "October_2024",
                                                                        "January_2025"))

ggplot(sym_summary, aes(x = Timepoint, y = mean_sym, color = Phenotype, group = Phenotype)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_sym - se_sym, ymax = mean_sym + se_sym, color = Phenotype),
                width = 0.3, linewidth = 0.8) +  # color matches points/lines
  facet_wrap(~ Species) +
  #facet_grid(rows = vars(Species)) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")
  ) +
  theme_classic(base_size = 16) +  # bigger base size
  labs(
    x = "",
    y = "Sym density (10^6 x cells/cm2)",
    color = "Phenotype"
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "italic", size = 16),
    #legend.position = c(0.85, 0.83),  # adjust x, y for placement
    legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 13, color = "black"),
    axis.text.y = element_text(size = 13, color = "black"),
    axis.title.y = element_text(size = 16)
  )
```

Values look slightly lower than I would expect...maybe due to the surface area being based on the fraggameter? Need the wax dipped SA. 

## Chlorophyll 

Read in data. 
```{r}
data_chl <- read.csv("../data/02-Physiology/Chlorophyll/chl_data.csv")
data_chl$ColonyID <- as.character(data_chl$ColonyID)
```

Merge raw data and metadata
```{r}
chl <- left_join(data_chl, metadata, by = "ColonyID")
```

Calculate means of technical replicates 
```{r}
chl <- chl %>%
  group_by(Date, ColonyID, Timepoint, Phenotype, Species) %>%
  dplyr::summarise(mean_A630 = mean(A630), # calculate means 
            mean_A663 = mean(A663), 
            mean_A750 = mean(A750))
```

Correct by blank
```{r}
chl_corrected <- chl %>%
  group_by(Date) %>%
  mutate(
    blank_750 = mean_A750[ColonyID == "Blank"][1],
    A630_corrected = mean_A630 - blank_750,
    A663_corrected = mean_A663 - blank_750) %>%
  ungroup()
```

Calculate chla and chlc2 values based on equations from Jeffrey and Humphrey 1975. units µg/ml
```{r}
chl_corrected <- chl_corrected %>%
    mutate(chla.ug.ml = (11.43 * A663_corrected)/0.6 - (0.64 * A630_corrected)/0.6,
           chlc2.ug.ml = (27.09 * A630_corrected)/0.6 - (3.63 * A663_corrected)/0.6) %>%
    filter(!is.na(Species)) # remove blanks from df
```

Plot data so far 
```{r}
chl_plot_data <- chl_corrected %>%
  filter(!is.na(Timepoint)) %>%
  mutate(Timepoint = factor(Timepoint, levels = c(
    "April_2023", 
    "September_2023", 
    "January_2024", 
    "May_2024",
    "October_2024"))) %>%
  pivot_longer(
    cols = c(chla.ug.ml, chlc2.ug.ml),
    names_to = "Chlorophyll_Type",
    values_to = "Concentration") %>%
  mutate(Chlorophyll_Type = case_when(
    Chlorophyll_Type == "chla.ug.ml" ~ "Chlorophyll a",
    Chlorophyll_Type == "chlc2.ug.ml" ~ "Chlorophyll c2"))

ggplot(chl_plot_data, aes(x = Timepoint, y = Concentration, fill = Phenotype)) +
  # Boxplot for data distribution summaries
  geom_boxplot(outlier.shape = NA, alpha = 0.6, position = position_dodge(0.8)) +
  # Overlay individual data points to see sample sizes/spread
  geom_point(aes(color = Phenotype), position = position_dodge(0.8), alpha = 0.6, size = 1.5) +
  # Facet by Species (rows) and Chlorophyll Type (columns)
  # scales = "free_y" allows Chl a and Chl c2 to have their own y-axis ranges
  facet_grid(Chlorophyll_Type ~ Species, scales = "free_y") +
  theme_bw() +
  labs(
    x = "Timepoint",
    y = "Concentration ($\\\\mu$g/mL)",
    title = "Chlorophyll Concentrations Across Timepoints",
    fill = "Phenotype",
    color = "Phenotype"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold", size = 10),
    legend.position = "bottom"
  )
```

Merge with surface area data and normalize 
```{r}
homog_vol_ml <- 10

chl_corrected <- chl_corrected %>%
  left_join(surface_area, by = c("ColonyID", "Species", "Phenotype", "Timepoint")) %>%
  mutate(chla.ug.cm2 = chla.ug.ml * homog_vol_ml / surface.area.cm2,
         chlc2.ug.cm2 = chlc2.ug.ml * homog_vol_ml / surface.area.cm2)
```

Plot normalized data
```{r}
chl_plot_data_norm <- chl_corrected %>%
  filter(!is.na(chla.ug.ml)) %>%
  mutate(Timepoint = factor(Timepoint, levels = c(
    "April_2023", 
    "September_2023", 
    "January_2024", 
    "May_2024",
    "October_2024"))) %>%
  pivot_longer(
    cols = c(chla.ug.cm2, chlc2.ug.cm2),
    names_to = "Chlorophyll_Type",
    values_to = "Concentration") %>%
  mutate(Chlorophyll_Type = case_when(
    Chlorophyll_Type == "chla.ug.cm2" ~ "Chlorophyll a",
    Chlorophyll_Type == "chlc2.ug.cm2" ~ "Chlorophyll c2")) #%>%
  #filter(!Timepoint == "April_2023")

ggplot(chl_plot_data_norm, aes(x = Timepoint, y = Concentration, fill = Phenotype)) +
  # Boxplot for data distribution summaries
  geom_boxplot(outlier.shape = NA, alpha = 0.6, position = position_dodge(0.8)) +
  # Overlay individual data points to see sample sizes/spread
  geom_point(aes(color = Phenotype), position = position_dodge(0.8), alpha = 0.6, size = 1.5) +
  # Facet by Species (rows) and Chlorophyll Type (columns)
  # scales = "free_y" allows Chl a and Chl c2 to have their own y-axis ranges
  facet_grid(Chlorophyll_Type ~ Species, scales = "free_y") +
  #scale_y_log10()+
  theme_bw() +
  labs(
    x = "Timepoint",
    title = "Chlorophyll Concentrations Across Timepoints",
    fill = "Phenotype",
    color = "Phenotype"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold", size = 10),
    legend.position = "bottom"
  )
```

Save as csv 
```{r}
write.csv(chl_corrected, "../output/02-Physiology/Chlorophyll/chlorophyll.csv")
```

Calculate mean, sem, and sd 
```{r}
chl_summary <- chl_corrected %>%
  group_by(Timepoint, Phenotype, Species) %>%
  summarise(
    mean_chla = mean(chla.ug.cm2, na.rm = TRUE),
    sd_chla = sd(chla.ug.cm2, na.rm = T),
    se_chla = sd(chla.ug.cm2, na.rm = T) / sqrt(length(chla.ug.cm2)),
    .groups = "drop") %>%
  drop_na() 
```

Plot! 
```{r}
# Set levels 
chl_summary$Timepoint <- factor(chl_summary$Timepoint, levels = c("April_2023",
                                                                        "September_2023",
                                                                        "January_2024", 
                                                                        "May_2024",
                                                                        "October_2024"))

ggplot(chl_summary, aes(x = Timepoint, y = mean_chla, color = Phenotype, group = Phenotype)) +
  #geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_chla - se_chla, ymax = mean_chla + se_chla, color = Phenotype),
                width = 0.3, linewidth = 0.8) +  # color matches points/lines
  facet_wrap(~ Species) +
  #facet_grid(rows = vars(Species)) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")
  ) +
  theme_classic(base_size = 16) +  # bigger base size
  labs(
    x = "",
    y = "Chlorophyll-a (ug/cm2)",
    color = "Phenotype"
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "italic", size = 16),
    #legend.position = c(0.85, 0.83),  # adjust x, y for placement
    legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 13, color = "black"),
    axis.text.y = element_text(size = 13, color = "black"),
    axis.title.y = element_text(size = 16)
  )
```

## AFDW

Read in AFDW data 
```{r}
afdw <- read.csv("../data/02-Physiology/AFDW/AFDW.csv")
afdw$ColonyID <- as.character(afdw$ColonyID)
```

Merge with metadata
```{r}
afdw <- afdw %>%
  left_join(metadata, by = "ColonyID")
```

Calculate biomass in mg/cm2
```{r}
homogenate_volume_ml <- 50

afdw_processed <- afdw %>%
  left_join(surface_area, by = c("ColonyID", "Timepoint")) %>%
  mutate(
    Dry_Tissue_Only_g = Dry_mass_g - Initial_mass_g,
    Ash_Only_g = Burnt_mass_g - Initial_mass_g,
    AFDW_per_mL_g = Dry_mass_g - Burnt_mass_g,
    Total_AFDW_g = AFDW_per_mL_g * homogenate_volume_ml,
    Total_AFDW_mg = (Total_AFDW_g * 1000),
    Tissue_Biomass_mg_cm2 = (Total_AFDW_g * 1000) / surface.area.cm2
  )
```

Plot data 
```{r}
afdw_processed <- afdw_processed %>%
  filter(!is.na(Tissue_Biomass_mg_cm2)) %>%
  mutate(Timepoint = factor(Timepoint, levels = c(
    "April_2023", 
    "September_2023", 
    "January_2024", 
    "May_2024",
    "October_2024", 
    "January_2025",
    "May_2025",
    "October_2025")))

ggplot(afdw_processed, aes(x = Timepoint, y = Tissue_Biomass_mg_cm2, fill = Phenotype, color = Phenotype)) +
  # Boxplot for data distribution summaries
  geom_boxplot(outlier.shape = NA, alpha = 0.6, position = position_dodge(0.8)) +
  # Overlay individual data points to see sample sizes/spread
  geom_point(aes(color = Phenotype), position = position_dodge(0.8), alpha = 0.6, size = 1.5) +
  # Facet by Species (rows) and Chlorophyll Type (columns)
  # scales = "free_y" allows Chl a and Chl c2 to have their own y-axis ranges
  facet_wrap(~Species, scales = "free_y") +
  scale_color_manual(values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  scale_fill_manual(values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  theme_bw() +
  labs(
    x = "Timepoint",
    y = "AFDW (mg/cm2)",
    fill = "Phenotype",
    color = "Phenotype"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold", size = 10),
    legend.position = "bottom"
  )
```

Save as csv 
```{r}
write.csv(afdw, "../output/02-Physiology/AFDW/AFDW_calculations.csv")
```

Calculate mean, sem, and sd 
```{r}
afdw_summary <- afdw %>%
  group_by(Timepoint, Phenotype, Species) %>%
  summarise(
    mean_afdw = mean(afdw_mg_cm2, na.rm = TRUE),
    sd_afdw = sd(afdw_mg_cm2, na.rm = T),
    se_afdw = sd(afdw_mg_cm2, na.rm = T) / sqrt(length(afdw_mg_cm2)),
    .groups = "drop") %>%
  drop_na() 
```

Plot! 
```{r}
# Set levels 
afdw_summary$Timepoint <- factor(afdw_summary$Timepoint, levels = c("April_2023",
                                                                        "September_2023",
                                                                        "January_2024", 
                                                                        "May_2024",
                                                                        "October_2024", 
                                                                        "January_2025", 
                                                                        "October_2025"))

ggplot(afdw_summary, aes(x = Timepoint, y = mean_afdw, color = Phenotype, group = Phenotype)) +
  #geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_afdw - se_afdw, ymax = mean_afdw + se_afdw, color = Phenotype),
                width = 0.3, linewidth = 0.8) +  # color matches points/lines
  facet_wrap(~ Species) +
  #facet_grid(rows = vars(Species)) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")
  ) +
  theme_classic(base_size = 16) +  # bigger base size
  labs(
    x = "",
    y = "Biomass (mg/cm2)",
    color = "Phenotype"
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "italic", size = 16),
    #legend.position = c(0.85, 0.83),  # adjust x, y for placement
    legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 13, color = "black"),
    axis.text.y = element_text(size = 13, color = "black"),
    axis.title.y = element_text(size = 16)
  )
```

## Protein & MAAs

Read in protein & MAA data
```{r}
prot <- read.csv("../data/Physiology/Protein_MAA_data.csv")
prot$ColonyID <- as.character(prot$ColonyID)
```

Read in surface area df and add time point col 
```{r}
surface_area <- read.csv("../data/Surface_Area/fragrammetry_SurfaceArea.csv")

surface_area <- surface_area %>%
  mutate(Timepoint = case_when(
    grepl("2023-04", Date) ~ "April_2023",
    grepl("2023-09", Date) ~ "September_2023",
    grepl("2024-01", Date) ~ "January_2024",
    grepl("2024-05", Date) ~ "May_2024",
    grepl("2024-10", Date) ~ "October_2024",
    grepl("2025-01", Date) ~ "January_2025",
    grepl("2025-05", Date) ~ "May_2025",
    grepl("2025-10", Date) ~ "October_2025",
    TRUE ~ NA_character_)) %>%
  rename(ColonyID = Colony_ID) %>%
  filter(Project == "CAREER - Phys")
```

Merge surface area and protein dfs and calculate protein/MAAs in mg/cm2
```{r}
homogenate_volume <- 50

prot <- prot %>%
  left_join(surface_area, by = c("ColonyID", "Timepoint")) %>%
  mutate(protein_mg_mL = (A235 - A280)/2.51,
         protein_mg = protein_mg_mL*homogenate_volume*Dilution,
         protein_mg_cm2 = protein_mg/Surface_Area_cm2,
         mycrosporine_glycine_mg_mL = A310/A280, 
         mycrosporine_glycine_mg = mycrosporine_glycine_mg_mL*homogenate_volume*Dilution,
         mycrosporine_glycine_mg_cm2 = mycrosporine_glycine_mg/Surface_Area_cm2,
         palythine_mg_mL = A320/A280,
         palythine_mg = palythine_mg_mL*homogenate_volume*Dilution,
         palythine_mg_cm2 = palythine_mg/Surface_Area_cm2,
         asterina_mg_mL = A330/A280,
         asterina_mg = asterina_mg_mL*homogenate_volume*Dilution,
         asterina_mg_cm2 = asterina_mg/Surface_Area_cm2,
         shinorine_mg_mL = A334/A280,
         shinorine_mg = shinorine_mg_mL*homogenate_volume*Dilution,
         shinorine_mg_cm2 = shinorine_mg/Surface_Area_cm2)
```

Save as csv 
XXXXXXXX

Calculate mean, sem, and sd 
```{r}
prot_summary <- prot %>%
  group_by(Timepoint, Phenotype, Species) %>%
  summarise(
    mean_prot = mean(protein_mg_cm2, na.rm = TRUE),
    sd_prot = sd(protein_mg_cm2, na.rm = T),
    se_prot = sd(protein_mg_cm2, na.rm = T) / sqrt(length(protein_mg_cm2)),
    mean_mycrosporine_glycine = mean(mycrosporine_glycine_mg_cm2, na.rm = TRUE),
    sd_mycrosporine_glycine = sd(mycrosporine_glycine_mg_cm2, na.rm = T),
    se_mycrosporine_glycine = sd(mycrosporine_glycine_mg_cm2, na.rm = T) / sqrt(length(mycrosporine_glycine_mg_cm2)),
    mean_palythine = mean(palythine_mg_cm2, na.rm = TRUE),
    sd_palythine = sd(palythine_mg_cm2, na.rm = T),
    se_palythine = sd(palythine_mg_cm2, na.rm = T) / sqrt(length(palythine_mg_cm2)),
    mean_asterina = mean(asterina_mg_cm2, na.rm = TRUE),
    sd_asterina = sd(asterina_mg_cm2, na.rm = T),
    se_asterina = sd(asterina_mg_cm2, na.rm = T) / sqrt(length(asterina_mg_cm2)),
    mean_shinorine = mean(shinorine_mg_cm2, na.rm = TRUE),
    sd_shinorine = sd(shinorine_mg_cm2, na.rm = T),
    se_shinorine = sd(shinorine_mg_cm2, na.rm = T) / sqrt(length(shinorine_mg_cm2)),
    .groups = "drop") %>%
  drop_na() 
```

Plot! 
```{r}
# Set levels 
prot_summary$Timepoint <- factor(prot_summary$Timepoint, levels = c("April_2023",
                                                                        "September_2023",
                                                                        "January_2024"))

# Protein
ggplot(prot_summary, aes(x = Timepoint, y = mean_prot, color = Phenotype, group = Phenotype)) +
  #geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_prot - se_prot, ymax = mean_prot + se_prot, color = Phenotype), width = 0.3, linewidth = 0.8) +  # color matches points/lines
  facet_wrap(~ Species) +
  #facet_grid(rows = vars(Species)) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  theme_classic(base_size = 16) +  # bigger base size
  labs(
    x = "",
    y = "Protein (mg/cm2)",
    color = "Phenotype") +
  theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "italic", size = 16),
    #legend.position = c(0.85, 0.83),  # adjust x, y for placement
    legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 13, color = "black"),
    axis.text.y = element_text(size = 13, color = "black"),
    axis.title.y = element_text(size = 16))

# Mycrosporine glycine
ggplot(prot_summary, aes(x = Timepoint, y = mean_mycrosporine_glycine, color = Phenotype, group = Phenotype)) +
  #geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_mycrosporine_glycine - se_mycrosporine_glycine, ymax = mean_mycrosporine_glycine + se_mycrosporine_glycine, color = Phenotype), width = 0.3, linewidth = 0.8) +  # color matches points/lines
  facet_wrap(~ Species) +
  #facet_grid(rows = vars(Species)) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  theme_classic(base_size = 16) +  # bigger base size
  labs(
    x = "",
    y = "Mycrosporine glycine (mg/cm2)",
    color = "Phenotype") +
  theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "italic", size = 16),
    #legend.position = c(0.85, 0.83),  # adjust x, y for placement
    legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 13, color = "black"),
    axis.text.y = element_text(size = 13, color = "black"),
    axis.title.y = element_text(size = 16))

# Palythine
ggplot(prot_summary, aes(x = Timepoint, y = mean_palythine, color = Phenotype, group = Phenotype)) +
  #geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_palythine - se_palythine, ymax = mean_palythine + se_palythine, color = Phenotype), width = 0.3, linewidth = 0.8) +  # color matches points/lines
  facet_wrap(~ Species) +
  #facet_grid(rows = vars(Species)) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  theme_classic(base_size = 16) +  # bigger base size
  labs(
    x = "",
    y = "Palythine (mg/cm2)",
    color = "Phenotype") +
  theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "italic", size = 16),
    #legend.position = c(0.85, 0.83),  # adjust x, y for placement
    legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 13, color = "black"),
    axis.text.y = element_text(size = 13, color = "black"),
    axis.title.y = element_text(size = 16))

# Asterina
ggplot(prot_summary, aes(x = Timepoint, y = mean_asterina, color = Phenotype, group = Phenotype)) +
  #geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_asterina - se_asterina, ymax = mean_asterina + se_asterina, color = Phenotype), width = 0.3, linewidth = 0.8) +  # color matches points/lines
  facet_wrap(~ Species) +
  #facet_grid(rows = vars(Species)) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  theme_classic(base_size = 16) +  # bigger base size
  labs(
    x = "",
    y = "Asterina (mg/cm2)",
    color = "Phenotype") +
  theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "italic", size = 16),
    #legend.position = c(0.85, 0.83),  # adjust x, y for placement
    legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 13, color = "black"),
    axis.text.y = element_text(size = 13, color = "black"),
    axis.title.y = element_text(size = 16))

# Shinorine
ggplot(prot_summary, aes(x = Timepoint, y = mean_shinorine, color = Phenotype, group = Phenotype)) +
  #geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_shinorine - se_shinorine, ymax = mean_shinorine + se_shinorine, color = Phenotype), width = 0.3, linewidth = 0.8) +  # color matches points/lines
  facet_wrap(~ Species) +
  #facet_grid(rows = vars(Species)) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  theme_classic(base_size = 16) +  # bigger base size
  labs(
    x = "",
    y = "Shinorine (mg/cm2)",
    color = "Phenotype") +
  theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "italic", size = 16),
    #legend.position = c(0.85, 0.83),  # adjust x, y for placement
    legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 13, color = "black"),
    axis.text.y = element_text(size = 13, color = "black"),
    axis.title.y = element_text(size = 16))
```

