---
title: "Physiology"
author: "Jill Ashey"
date: "2025-11-24"
output: html_document
---

Script for all physiology (sym counts, chlorophyll, protein, MAAs, and AFDW) analyses. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(purrr)
library(readr)
library(stringr)
library(emmeans)
```

## Prep for analyses

Run this before running any ANOVAs
```{r}
options(contrasts = c("contr.sum", "contr.poly"))
```

Read in colony metadata
```{r}
metadata <- read.csv("../data/02-Physiology/Pairs_metadata.csv")
metadata$ColonyID <- as.character(metadata$ColonyID)
```

Read in wax surface area data 
```{r}
surface_area <- read.csv("../output/02-Physiology/SurfaceArea/SurfaceArea_wax_calculations.csv")
surface_area$ColonyID <- as.character(surface_area$ColonyID)
```

## Fv/Fm

### Process data 

Read in PAM data
```{r}
pam <- read.csv("../data/02-Physiology/PAM/PAM.csv")
pam$ColonyID <- as.character(pam$ColonyID)
```

Join with colony metadata
```{r}
pam <- pam %>%
  left_join(metadata, by = "ColonyID")
```

Calculate mean, sem, and sd 
```{r}
pam_summary <- pam %>%
  group_by(Timepoint, Phenotype, Species) %>%
  summarise(
    mean_fvfm = mean(Fv_Fm, na.rm = TRUE),
    sd_fvfm = sd(Fv_Fm, na.rm = T),
    se_fvfm = sd(Fv_Fm, na.rm = T) / sqrt(length(Fv_Fm)),
    .groups = "drop") %>%
  drop_na()# %>%
  #filter(Light == "Light")
  #filter(!Timepoint == "Jan25") # remove Jan25 for now 
```

Plot data
```{r}
pam_summary$Timepoint <- gsub("_", " ", pam_summary$Timepoint)

pam_summary <- pam_summary %>%
  #filter(!is.na(Fv_Fm)) %>%
  mutate(Timepoint = factor(Timepoint, levels = c(
    "April 2023", 
    "September 2023", 
    "January 2024", 
    "May 2024",
    "October 2024", 
    "January 2025",
    "May 2025",
    "October 2025",
    "January 2026")))

y_label <- expression(bold(F[v]/F[m]))
pam_plot <- ggplot(pam_summary, aes(x = Timepoint, y = mean_fvfm, color = Phenotype, group = Phenotype)) +
  #geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +
  geom_line(linewidth = 2.2) +
  geom_point(size = 7) +
  geom_errorbar(aes(ymin = mean_fvfm - se_fvfm, ymax = mean_fvfm + se_fvfm, color = Phenotype),
                width = 0.4, linewidth = 1.8) +  # color matches points/lines
  #facet_wrap(~ Species + Light) +
  facet_grid(cols = vars(Species)) +
  scale_y_continuous(breaks = seq(0.4, 0.8, by = 0.05)) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")
  ) +
  theme_classic() +  
  labs(
    x = "",
    y = y_label,
    color = "Phenotype"
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "bold.italic", size = 24),
    #legend.position = c(0.85, 0.83),  # adjust x, y for placement
    legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 18, face = "bold"),
    legend.text = element_text(size = 17),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 20, color = "black"),
    axis.text.y = element_text(size = 24, color = "black"),
    axis.title.y = element_text(size = 26, face = "bold")
); pam_plot

ggsave("../output/02-Physiology/PAM/PAM_KBay.png", 
       plot = pam_plot, width = 20, height = 8, units = "in", dpi = 300)
ggsave("../output/02-Physiology/PAM/PAM_KBay.pdf", 
       plot = pam_plot, width = 20, height = 8, units = "in", device = "pdf", dpi = 300)
```

### Stats

#### Mcap

Subset by Mcap only 
```{r}
pam_mcap <- pam %>%
  filter(Species == "Montipora capitata")
```

Check for normality
```{r}
hist(pam_mcap$Fv_Fm)

# Q-Q Plot
qqnorm(pam_mcap$Fv_Fm)
qqline(pam_mcap$Fv_Fm, col = "red")

# Shapiro-Wilk Test
shapiro.test(pam_mcap$Fv_Fm)
```

Normal enough.

Run linear mixed effects model 
```{r}
pam.mcap.model <- lme4::lmer(Fv_Fm ~ Timepoint * Phenotype + (1|ColonyID), data = pam_mcap)
car::Anova(pam.mcap.model, type = 3)

# Model normality 
par(mfrow = c(1, 2)) 
plot(pam.mcap.model, which = 1) 
lattice::qqmath(pam.mcap.model)

# Post hoc tests 
emmeans(pam.mcap.model, pairwise ~ Timepoint)

# Check plot to make sure results make sense
ggplot(pam_mcap, aes(x = Timepoint, y = Fv_Fm, color = Phenotype, group = Phenotype, shape = PAM_type)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  facet_wrap(~Species) +
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

#### Pcomp

Subset by Pcomp only 
```{r}
pam_pcomp <- pam %>%
  filter(Species == "Porites compressa")
```

Check for normality
```{r}
hist(pam_pcomp$Fv_Fm)

# Q-Q Plot
qqnorm(pam_pcomp$Fv_Fm)
qqline(pam_pcomp$Fv_Fm, col = "red")

# Shapiro-Wilk Test
shapiro.test(pam_pcomp$Fv_Fm)
```

Normal.

Run linear mixed effects model 
```{r}
pam.pcomp.model <- lme4::lmer(Fv_Fm ~ Timepoint * Phenotype + (1|ColonyID), data = pam_pcomp)
car::Anova(pam.pcomp.model, type = 3)

# Model normality 
par(mfrow = c(1, 2)) 
plot(pam.pcomp.model, which = 1) 
lattice::qqmath(pam.pcomp.model)

# Post hoc tests 
emmeans(pam.pcomp.model, pairwise ~ Timepoint)
emmeans(pam.pcomp.model, pairwise ~ Phenotype)
emmeans(pam.pcomp.model, pairwise ~ Timepoint | Phenotype)
emmeans(pam.pcomp.model, pairwise ~ Phenotype | Timepoint)

# Check plot to make sure results make sense
ggplot(pam_pcomp, aes(x = Timepoint, y = Fv_Fm, color = Phenotype, group = Phenotype, shape = PAM_type)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  facet_wrap(~Species) +
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

## Sym counts 

Read in metadata 
```{r}
meta_sym <- read.csv("../data/02-Physiology/Symbiont_density/sym_count_metadata.csv")

# Remove extra spaces 
meta_sym <- meta_sym %>%
  mutate(
    Run = str_replace_all(Run, "\\s+", ""),  # "Run 1" -> "Run1"
    Well = str_trim(Well),
    Date = as.character(Date),
    ColonyID = as.character(ColonyID))
```

Read in raw data from guava
```{r}
files <- list.files(path = "../data/02-Physiology/Symbiont_density/raw_guava", pattern = "*.csv", full.names = TRUE)

read_with_metadata <- function(f) {
  df <- read.csv(f, skip = 4, header = TRUE, stringsAsFactors = FALSE)
  
  # extract filename without path and extension
  filename <- basename(f)
  filename <- str_remove(filename, "\\.csv$")

  # split filename at underscore, e.g., "20230805_Run1" -> c("20230805", "Run1")
  parts <- str_split(filename, "_", simplify = TRUE)

  # add columns for Date and Run parsed from filename
  df %>%
    mutate(
      Date = parts[1],
      Run = parts[2],
      # ensure consistent types and clean whitespace
      Well = str_trim(Well),
      Sample.ID = as.character(Sample.ID)
    )
}

all_dfs <- lapply(files, read_with_metadata)

combined_df <- bind_rows(all_dfs)

combined_df <- combined_df %>%
  mutate(
    Run = str_replace_all(Run, "\\s+", ""),
    Well = str_trim(Well),
    Date = as.character(Date)  # ensure same type
  )
```

Merge raw data and guava metadata. Merge w/ colony metadata as well
```{r}
sym_counts <- left_join(meta_sym, combined_df, by = c("Date", "Run", "Well"))

sym_counts <- sym_counts %>%
  left_join(metadata, by = "ColonyID")
```

Remove samples marked "Do not use"
```{r}
sym_counts <- sym_counts %>%
  filter(is.na(Use) | Use == "")
```

Calculate mean counts per sample and number of cells per sample 
```{r}
sym_counts_avg <- sym_counts %>%
  mutate(across(starts_with("CHLOROPHYLL.POS.Concentration"), ~ as.numeric(as.character(.)))) %>%
  filter(!is.na(pick(starts_with("CHLOROPHYLL.POS.Concentration"))[[1]])) %>%
  group_by(Date, Run, Timepoint, ColonyID, Species, Phenotype, Pellet_Resuspend_Vol_mL) %>%
  summarise(
    mean_cells_mL = mean(pick(starts_with("CHLOROPHYLL.POS.Concentration"))[[1]], na.rm = TRUE),
    cells = mean_cells_mL * unique(Pellet_Resuspend_Vol_mL),
    .groups = "drop") %>%
  full_join(surface_area, by = c("ColonyID", "Timepoint")) %>%
  mutate(surface.area.cm2 = ifelse(is.na(surface.area.cm2), 10, surface.area.cm2)) %>%
  mutate(cells_cm2 = cells/surface.area.cm2,
          cells_cm2_M = cells_cm2 / 1e6) %>%
  filter(ColonyID != "blank" & !is.na(ColonyID)) %>%
  rename(Phenotype = Phenotype.x, Species = Species.x) %>%
  select(Timepoint, ColonyID, Species, Phenotype, cells_cm2, cells_cm2_M)
```

Some samples were manually counted when the guava was down. Read in and process manual data 
```{r}
# Read in data 
sym_manual <- read.csv("../data/02-Physiology/Symbiont_density/manual_sym_counts.csv", header = T)
sym_manual$ColonyID <- as.character(sym_manual$ColonyID)

sym_manual_calc <- sym_manual %>%
  mutate(Dilution = ifelse(is.na(Dilution), 1, Dilution)) %>%
  dplyr::select(ColonyID, Timepoint, Squares.Counted, Dilution, matches("Count[0-9]")) %>%
  pivot_longer(cols = starts_with("Count"), names_to = "rep", values_to = "val") %>%
  group_by(ColonyID, Timepoint, Squares.Counted, Dilution) %>%
  dplyr::summarise(mean_val = mean(val, na.rm = TRUE), .groups = "drop") %>%
  left_join(surface_area, by = c("ColonyID", "Timepoint")) %>%
  mutate(
    # 1. Cells per mL in the counting tube
    cells.mL = (mean_val / Squares.Counted) * 10000 * Dilution,
    
    # 2. Total cells in the original 10mL slurry
    # If the manual count was done on the same 1mL aliquot resuspended in 0.9mL SDS:
    cells = cells.mL * 0.9 * (10 / 1.0), 
    
    cells_cm2 = cells / surface.area.cm2,
    cells_cm2_M = cells_cm2 / 1e6) %>%
  select(Timepoint, ColonyID, Species, Phenotype, cells_cm2, cells_cm2_M)
```

The following October 2025 samples need to be removed from the `sym_counts_avg` guava counts, as we will be using the hemocytometer counts for them: 202, 244, 2, 41, 240, 248, 239, 28, 225, 209, 236, 20, 243, 235, 205, 3, 237, 206, 204, 1, 247.

Filter the above samples from the guava df 
```{r}
to_remove <- c("202", "244", "2", "41", "240", "248", "239", "28", "225", 
               "209", "236", "20", "243", "235", "205", "3", "237", 
               "206", "204", "1", "247")

sym_counts_avg_filt <- sym_counts_avg %>%
  filter(!(Timepoint == "October_2025" & ColonyID %in% to_remove))
```

Bind dfs together
```{r}
sym_data <- rbind(sym_counts_avg_filt, sym_manual_calc)
```

Save as csv 
```{r}
write.csv(sym_data, "../output/02-Physiology/Symbiont_density/sym_density.csv")
```

Calculate mean, sem, and sd 
```{r}
sym_summary <- sym_data %>%
  group_by(Timepoint, Phenotype, Species) %>%
  summarise(
    mean_sym = mean(cells_cm2_M, na.rm = TRUE),
    sd_sym = sd(cells_cm2_M, na.rm = T),
    se_sym = sd(cells_cm2_M, na.rm = T) / sqrt(length(cells_cm2_M)),
    .groups = "drop") %>%
  drop_na()
```

Plot!
```{r}
sym_summary$Timepoint <- gsub("_", " ", sym_summary$Timepoint)

# Set levels 
sym_summary$Timepoint <- factor(sym_summary$Timepoint, levels = c("April 2023",
                                                                        "September 2023",
                                                                        "January 2024", 
                                                                        "May 2024", 
                                                                        "October 2024", 
                                                                        "January 2025", 
                                                                        "May 2025",
                                                                        "October 2025"))

y_label <- expression(bold(paste("Symbiont density (", 10^6, " cells ", cm^-2, ")")))
sym_plot <- ggplot(sym_summary, aes(x = Timepoint, y = mean_sym, color = Phenotype, group = Phenotype)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +
  geom_line(linewidth = 2.2) +
  geom_point(size = 7) +
  geom_errorbar(aes(ymin = mean_sym - se_sym, ymax = mean_sym + se_sym, color = Phenotype),
                width = 0.4, linewidth = 1.8) +  # color matches points/lines
  #facet_wrap(~ Species + Light) +
  facet_grid(cols = vars(Species)) +
  scale_y_continuous(breaks = seq(-2, 10, by = 0.2)) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")
  ) +
  theme_classic() +  
  labs(
    x = "",
    y = y_label,
    color = "Phenotype"
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "bold.italic", size = 24),
    #legend.position = c(0.85, 0.83),  # adjust x, y for placement
    legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 18, face = "bold"),
    legend.text = element_text(size = 17),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 20, color = "black"),
    axis.text.y = element_text(size = 24, color = "black"),
    axis.title.y = element_text(size = 26, face = "bold")
); sym_plot

ggsave("../output/02-Physiology/Symbiont_density/Symbiont_density_KBay.png", 
       plot = sym_plot, width = 20, height = 8, units = "in", dpi = 300)
ggsave("../output/02-Physiology/Symbiont_density/Symbiont_density_KBay.pdf", 
       plot = sym_plot, width = 20, height = 8, units = "in", device = "pdf", dpi = 300)
```

Plot looks really weird...need to check the math and data. 


## Chlorophyll 

Read in data. 
```{r}
data_chl <- read.csv("../data/02-Physiology/Chlorophyll/chl_data.csv")
data_chl$ColonyID <- as.character(data_chl$ColonyID)
```

Merge raw data and metadata
```{r}
chl <- left_join(data_chl, metadata, by = "ColonyID")
```

Calculate means of technical replicates 
```{r}
chl <- chl %>%
  group_by(Date, ColonyID, Timepoint, Phenotype, Species) %>%
  dplyr::summarise(mean_A630 = mean(A630), # calculate means 
            mean_A663 = mean(A663), 
            mean_A750 = mean(A750))
```

Correct by blank
```{r}
chl_corrected <- chl %>%
  group_by(Date) %>%
  mutate(
    blank_750 = mean_A750[ColonyID == "Blank"][1],
    A630_corrected = mean_A630 - blank_750,
    A663_corrected = mean_A663 - blank_750) %>%
  ungroup()
```

Calculate chla and chlc2 values based on equations from Jeffrey and Humphrey 1975. units µg/ml
```{r}
chl_corrected <- chl_corrected %>%
    mutate(chla.ug.ml = (11.43 * A663_corrected)/0.6 - (0.64 * A630_corrected)/0.6,
           chlc2.ug.ml = (27.09 * A630_corrected)/0.6 - (3.63 * A663_corrected)/0.6) %>%
    filter(!is.na(Species)) # remove blanks from df
```

Plot data so far 
```{r}
chl_plot_data <- chl_corrected %>%
  filter(!is.na(Timepoint)) %>%
  mutate(Timepoint = factor(Timepoint, levels = c(
    "April_2023", 
    "September_2023", 
    "January_2024", 
    "May_2024",
    "October_2024",
    "January_2025"))) %>%
  pivot_longer(
    cols = c(chla.ug.ml, chlc2.ug.ml),
    names_to = "Chlorophyll_Type",
    values_to = "Concentration") %>%
  mutate(Chlorophyll_Type = case_when(
    Chlorophyll_Type == "chla.ug.ml" ~ "Chlorophyll a",
    Chlorophyll_Type == "chlc2.ug.ml" ~ "Chlorophyll c2"))

ggplot(chl_plot_data, aes(x = Timepoint, y = Concentration, fill = Phenotype)) +
  # Boxplot for data distribution summaries
  geom_boxplot(outlier.shape = NA, alpha = 0.6, position = position_dodge(0.8)) +
  # Overlay individual data points to see sample sizes/spread
  geom_point(aes(color = Phenotype), position = position_dodge(0.8), alpha = 0.6, size = 1.5) +
  # Facet by Species (rows) and Chlorophyll Type (columns)
  # scales = "free_y" allows Chl a and Chl c2 to have their own y-axis ranges
  facet_grid(Chlorophyll_Type ~ Species, scales = "free_y") +
  theme_bw() +
  labs(
    x = "Timepoint",
    y = "Concentration ($\\\\mu$g/mL)",
    title = "Chlorophyll Concentrations Across Timepoints",
    fill = "Phenotype",
    color = "Phenotype"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold", size = 10),
    legend.position = "bottom"
  )
```

Merge with surface area data and normalize 
```{r}
homog_vol_ml <- 10

chl_corrected <- chl_corrected %>%
  left_join(surface_area, by = c("ColonyID", "Species", "Phenotype", "Timepoint")) %>%
  mutate(chla.ug.cm2 = chla.ug.ml * homog_vol_ml / surface.area.cm2,
         chlc2.ug.cm2 = chlc2.ug.ml * homog_vol_ml / surface.area.cm2)
```

Plot normalized data
```{r}
chl_plot_data_norm <- chl_corrected %>%
  filter(!is.na(chla.ug.ml)) %>%
  mutate(Timepoint = factor(Timepoint, levels = c(
    "April_2023", 
    "September_2023", 
    "January_2024", 
    "May_2024",
    "October_2024", 
    "January_2025"))) %>%
  pivot_longer(
    cols = c(chla.ug.cm2, chlc2.ug.cm2),
    names_to = "Chlorophyll_Type",
    values_to = "Concentration") %>%
  mutate(Chlorophyll_Type = case_when(
    Chlorophyll_Type == "chla.ug.cm2" ~ "Chlorophyll a",
    Chlorophyll_Type == "chlc2.ug.cm2" ~ "Chlorophyll c2")) #%>%
  #filter(!Timepoint == "April_2023")

ggplot(chl_plot_data_norm, aes(x = Timepoint, y = Concentration, fill = Phenotype)) +
  # Boxplot for data distribution summaries
  geom_boxplot(outlier.shape = NA, alpha = 0.6, position = position_dodge(0.8)) +
  # Overlay individual data points to see sample sizes/spread
  geom_point(aes(color = Phenotype), position = position_dodge(0.8), alpha = 0.6, size = 1.5) +
  # Facet by Species (rows) and Chlorophyll Type (columns)
  # scales = "free_y" allows Chl a and Chl c2 to have their own y-axis ranges
  facet_grid(Chlorophyll_Type ~ Species, scales = "free_y") +
  #scale_y_log10()+
  theme_bw() +
  labs(
    x = "Timepoint",
    title = "Chlorophyll Concentrations Across Timepoints",
    fill = "Phenotype",
    color = "Phenotype"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold", size = 10),
    legend.position = "bottom"
  )
```

Save as csv 
```{r}
write.csv(chl_corrected, "../output/02-Physiology/Chlorophyll/chlorophyll.csv")
```

Calculate mean, sem, and sd 
```{r}
chl_summary <- chl_corrected %>%
  group_by(Timepoint, Phenotype, Species) %>%
  summarise(
    mean_chla = mean(chla.ug.cm2, na.rm = TRUE),
    sd_chla = sd(chla.ug.cm2, na.rm = T),
    se_chla = sd(chla.ug.cm2, na.rm = T) / sqrt(length(chla.ug.cm2)),
    .groups = "drop") %>%
  drop_na() 
```

Plot! 
```{r}
chl_summary$Timepoint <- gsub("_", " ", chl_summary$Timepoint)

# Set levels 
chl_summary$Timepoint <- factor(chl_summary$Timepoint, levels = c("April 2023",
                                                                        "September 2023",
                                                                        "January 2024", 
                                                                        "May 2024", 
                                                                        "October 2024", 
                                                                        "January 2025"))

y_label <- expression(bold(paste("Chlorophyll-a (", mu, "g ", cm^-2, ")")))
chl_plot <- ggplot(chl_summary, aes(x = Timepoint, y = mean_chla, color = Phenotype, group = Phenotype)) +
  #geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +
  geom_line(linewidth = 2.2) +
  geom_point(size = 7) +
  geom_errorbar(aes(ymin = mean_chla - se_chla, ymax = mean_chla + se_chla, color = Phenotype),
                width = 0.4, linewidth = 1.8) +  # color matches points/lines
  #facet_wrap(~ Species + Light) +
  facet_grid(cols = vars(Species)) +
  scale_y_continuous(breaks = seq(-2, 30, by = 5)) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")
  ) +
  theme_classic() +  
  labs(
    x = "",
    y = y_label,
    color = "Phenotype"
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "bold.italic", size = 24),
    #legend.position = c(0.85, 0.83),  # adjust x, y for placement
    legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 18, face = "bold"),
    legend.text = element_text(size = 17),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 20, color = "black"),
    axis.text.y = element_text(size = 24, color = "black"),
    axis.title.y = element_text(size = 26, face = "bold")
); chl_plot

ggsave("../output/02-Physiology/Chlorophyll/Chlorophyll_KBay.png", 
       plot = chl_plot, width = 18, height = 8, units = "in", dpi = 300)
ggsave("../output/02-Physiology/Chlorophyll/Chlorophyll_KBay.pdf", 
       plot = chl_plot, width = 18, height = 8, units = "in", device = "pdf", dpi = 300)
```

## AFDW

Read in AFDW data 
```{r}
afdw <- read.csv("../data/02-Physiology/AFDW/AFDW.csv")
afdw$ColonyID <- as.character(afdw$ColonyID)
```

Merge with metadata
```{r}
afdw <- afdw %>%
  left_join(metadata, by = "ColonyID")
```

Calculate biomass in mg/cm2
```{r}
homogenate_volume_ml <- 50

afdw_processed <- afdw %>%
  left_join(surface_area, by = c("ColonyID", "Timepoint")) %>%
  mutate(
    Dry_Tissue_Only_g = Dry_mass_g - Initial_mass_g,
    Ash_Only_g = Burnt_mass_g - Initial_mass_g,
    AFDW_per_mL_g = Dry_mass_g - Burnt_mass_g,
    Total_AFDW_g = AFDW_per_mL_g * homogenate_volume_ml,
    Total_AFDW_mg = (Total_AFDW_g * 1000),
    Tissue_Biomass_mg_cm2 = (Total_AFDW_g * 1000) / surface.area.cm2) %>%
  rename(Phenotype = Phenotype.x, Species = Species.x) %>%
  select(Timepoint, ColonyID, Species, Phenotype, Tissue_Biomass_mg_cm2)
```

Plot data 
```{r}
afdw_processed <- afdw_processed %>%
  filter(!is.na(Tissue_Biomass_mg_cm2)) %>%
  mutate(Timepoint = factor(Timepoint, levels = c(
    "April_2023", 
    "September_2023", 
    "January_2024", 
    "May_2024",
    "October_2024", 
    "January_2025",
    "May_2025",
    "October_2025")))

ggplot(afdw_processed, aes(x = Timepoint, y = Tissue_Biomass_mg_cm2, fill = Phenotype, color = Phenotype)) +
  # Boxplot for data distribution summaries
  geom_boxplot(outlier.shape = NA, alpha = 0.6, position = position_dodge(0.8)) +
  # Overlay individual data points to see sample sizes/spread
  geom_point(aes(color = Phenotype), position = position_dodge(0.8), alpha = 0.6, size = 1.5) +
  # Facet by Species (rows) and Chlorophyll Type (columns)
  # scales = "free_y" allows Chl a and Chl c2 to have their own y-axis ranges
  facet_wrap(~Species, scales = "free_y") +
  scale_color_manual(values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  scale_fill_manual(values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  theme_bw() +
  labs(
    x = "Timepoint",
    y = "AFDW (mg/cm2)",
    fill = "Phenotype",
    color = "Phenotype"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold", size = 10),
    legend.position = "bottom"
  )
```

Save as csv 
```{r}
write.csv(afdw_processed, "../output/02-Physiology/AFDW/AFDW.csv")
```

Calculate mean, sem, and sd 
```{r}
afdw_summary <- afdw_processed %>%
  group_by(Timepoint, Phenotype, Species) %>%
  summarise(
    mean_afdw = mean(Tissue_Biomass_mg_cm2, na.rm = TRUE),
    sd_afdw = sd(Tissue_Biomass_mg_cm2, na.rm = T),
    se_afdw = sd(Tissue_Biomass_mg_cm2, na.rm = T) / sqrt(length(Tissue_Biomass_mg_cm2)),
    .groups = "drop") %>%
  drop_na() %>%
  filter(!Timepoint == "May_2025")
```

Plot! 
```{r}
afdw_summary$Timepoint <- gsub("_", " ", afdw_summary$Timepoint)

# Set levels 
afdw_summary$Timepoint <- factor(afdw_summary$Timepoint, levels = c("April 2023",
                                                                        "September 2023",
                                                                        "January 2024", 
                                                                        "May 2024", 
                                                                        "October 2024", 
                                                                        "January 2025", 
                                                                        "October 2025"))

y_label <- expression(bold(paste("Biomass (mg ",cm^-2, ")")))
afdw_plot <- ggplot(afdw_summary, aes(x = Timepoint, y = mean_afdw, color = Phenotype, group = Phenotype)) +
  geom_line(linewidth = 2.2) +
  geom_point(size = 7) +
  geom_errorbar(aes(ymin = mean_afdw - se_afdw, ymax = mean_afdw + se_afdw, color = Phenotype),
                width = 0.4, linewidth = 1.8) +  # color matches points/lines
  #facet_wrap(~ Species + Light) +
  facet_grid(cols = vars(Species)) +
  scale_y_continuous(breaks = seq(0, 10, by = 2)) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")
  ) +
  theme_classic() +  
  labs(
    x = "",
    y = y_label,
    color = "Phenotype"
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "bold.italic", size = 24),
    #legend.position = c(0.85, 0.83),  # adjust x, y for placement
    legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 18, face = "bold"),
    legend.text = element_text(size = 17),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 20, color = "black"),
    axis.text.y = element_text(size = 24, color = "black"),
    axis.title.y = element_text(size = 26, face = "bold")
); afdw_plot

ggsave("../output/02-Physiology/AFDW/AFDW_KBay.png", 
       plot = afdw_plot, width = 20, height = 8, units = "in", dpi = 300)
ggsave("../output/02-Physiology/AFDW/AFDW_KBay.pdf", 
       plot = afdw_plot, width = 20, height = 8, units = "in", device = "pdf", dpi = 300)
```

## Protein & MAAs

Read in protein & MAA data
```{r}
data_prot <- read.csv("../data/02-Physiology/Protein/Protein_MAA_data.csv")
data_prot$ColonyID <- as.character(data_prot$ColonyID)
```

Merge raw data and metadata
```{r}
prot <- left_join(data_prot, metadata, by = "ColonyID")
```

Calculate means of technical replicates 
```{r}
prot <- prot %>%
  group_by(Date, ColonyID, Timepoint, Phenotype, Species, Dilution) %>%
  dplyr::summarise(mean_A235 = mean(A235), # calculate means 
            mean_A280 = mean(A280), 
            mean_A310 = mean(A310),
            mean_A320 = mean(A320),
            mean_A330 = mean(A330),
            mean_A334 = mean(A334))
```

Calculate protein and MAAs in mg/mL using the equation derived from Whitaker and Granum 1980. Idk where the MAAs calculation came from units mg/mL
```{r}
prot <- prot %>%
    mutate(prot.mg.ml = (mean_A235 - mean_A280)/2.51,
           mycro.mg.ml = (mean_A310/mean_A280),
           paly.mg.ml = (mean_A320/mean_A280),
           aster.mg.ml = (mean_A330/mean_A280),
           shin.mg.ml = (mean_A334/mean_A280)) %>%
    filter(!is.na(Species)) # remove blanks from df
```

Plot data so far 
```{r}
prot_plot_data <- prot %>%
  filter(!is.na(Timepoint)) %>%
  mutate(Timepoint = factor(Timepoint, levels = c(
    "April_2023", 
    "September_2023", 
    "January_2024", 
    "May_2024",
    "October_2024"))) %>%
  pivot_longer(
    cols = c(prot.mg.ml, mycro.mg.ml, paly.mg.ml, aster.mg.ml, shin.mg.ml),
    names_to = "Protein_Type",
    values_to = "Concentration") %>%
  mutate(Chlorophyll_Type = case_when(
    Protein_Type == "prot.mg.ml" ~ "Protein",
    Protein_Type == "mycro.mg.ml" ~ "Mycrosporine",
    Protein_Type == "paly.mg.ml" ~ "Palythine",
    Protein_Type == "aster.mg.ml" ~ "Asterina",
    Protein_Type == "shin.mg.ml" ~ "Shinorine"))

ggplot(prot_plot_data, aes(x = Timepoint, y = Concentration, fill = Phenotype)) +
  # Boxplot for data distribution summaries
  geom_boxplot(outlier.shape = NA, alpha = 0.6, position = position_dodge(0.8)) +
  # Overlay individual data points to see sample sizes/spread
  geom_point(aes(color = Phenotype), position = position_dodge(0.8), alpha = 0.6, size = 1.5) +
  # Facet by Species (rows) and Chlorophyll Type (columns)
  # scales = "free_y" allows Chl a and Chl c2 to have their own y-axis ranges
  facet_grid(Protein_Type ~ Species, scales = "free_y") +
  theme_bw() +
  labs(
    x = "Timepoint",
    y = "Concentration (mg/mL)",
    title = "Protein/MAAs Across Timepoints",
    fill = "Phenotype",
    color = "Phenotype"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold", size = 10),
    legend.position = "bottom"
  )
```

Merge with surface area data and normalize 
```{r}
homog_vol_ml <- 50

prot <- prot %>%
  left_join(surface_area, by = c("ColonyID", "Species", "Phenotype", "Timepoint")) %>%
  mutate(prot.mg.cm2 = (homog_vol_ml * prot.mg.ml * Dilution)/surface.area.cm2,
         mycro.mg.cm2 = (homog_vol_ml * mycro.mg.ml * Dilution)/surface.area.cm2,
         paly.mg.cm2 = (homog_vol_ml * paly.mg.ml * Dilution)/surface.area.cm2, 
         aster.mg.cm2 = (homog_vol_ml * aster.mg.ml * Dilution)/surface.area.cm2, 
         shin.mg.cm2 = (homog_vol_ml * shin.mg.ml * Dilution)/surface.area.cm2)
```

Plot normalized data 
```{r}
prot_plot_data_norm <- prot %>%
  filter(!is.na(Timepoint)) %>%
  mutate(Timepoint = factor(Timepoint, levels = c(
    "April_2023", 
    "September_2023", 
    "January_2024", 
    "May_2024",
    "October_2024"))) %>%
  pivot_longer(
    cols = c(prot.mg.cm2, mycro.mg.cm2, paly.mg.cm2, aster.mg.cm2, shin.mg.cm2),
    names_to = "Protein_Type",
    values_to = "Concentration") %>%
  mutate(Chlorophyll_Type = case_when(
    Protein_Type == "prot.mg.cm2" ~ "Protein",
    Protein_Type == "mycro.mg.cm2" ~ "Mycrosporine",
    Protein_Type == "paly.mg.cm2" ~ "Palythine",
    Protein_Type == "aster.mg.cm2" ~ "Asterina",
    Protein_Type == "shin.mg.cm2" ~ "Shinorine"))

ggplot(prot_plot_data_norm, aes(x = Timepoint, y = Concentration, fill = Phenotype)) +
  # Boxplot for data distribution summaries
  geom_boxplot(outlier.shape = NA, alpha = 0.6, position = position_dodge(0.8)) +
  # Overlay individual data points to see sample sizes/spread
  geom_point(aes(color = Phenotype), position = position_dodge(0.8), alpha = 0.6, size = 1.5) +
  # Facet by Species (rows) and Chlorophyll Type (columns)
  # scales = "free_y" allows Chl a and Chl c2 to have their own y-axis ranges
  facet_grid(Protein_Type ~ Species, scales = "free_y") +
  #scale_y_log10()+
  theme_bw() +
  labs(
    x = "Timepoint",
    title = "Protein/MAAs Across Timepoints",
    fill = "Phenotype",
    color = "Phenotype"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold", size = 10),
    legend.position = "bottom"
  )
```

243 in April 2023 looks like an outlier (much higher values that other samples). Maybe remove for formal analysis? 

Save as csv 
```{r}
write.csv(prot, "../output/02-Physiology/Protein/protein.csv")
```

Calculate mean, sem, and sd of protein only 
```{r}
prot_summary <- prot %>%
  group_by(Timepoint, Phenotype, Species) %>%
  summarise(
    mean_prot = mean(prot.mg.cm2, na.rm = TRUE),
    sd_prot = sd(prot.mg.cm2, na.rm = T),
    se_prot = sd(prot.mg.cm2, na.rm = T) / sqrt(length(prot.mg.cm2)),
    .groups = "drop") %>%
  drop_na() 
```

Plot! 
```{r}
prot_summary$Timepoint <- gsub("_", " ", prot_summary$Timepoint)

# Set levels 
prot_summary$Timepoint <- factor(prot_summary$Timepoint, levels = c("April 2023",
                                                                        "September 2023",
                                                                        "January 2024", 
                                                                        "May 2024", 
                                                                        "October 2024"))

y_label <- expression(bold(paste("Protein (mg ",cm^-2, ")")))
prot_plot <- ggplot(prot_summary, aes(x = Timepoint, y = mean_prot, color = Phenotype, group = Phenotype)) +
  geom_line(linewidth = 2.2) +
  geom_point(size = 7) +
  geom_errorbar(aes(ymin = mean_prot - se_prot, ymax = mean_prot + se_prot, color = Phenotype),
                width = 0.4, linewidth = 1.8) +  # color matches points/lines
  #facet_wrap(~ Species + Light) +
  facet_grid(cols = vars(Species)) +
  scale_y_continuous(breaks = seq(0, 10, by = 2)) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")
  ) +
  theme_classic() +  
  labs(
    x = "",
    y = y_label,
    color = "Phenotype"
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "bold.italic", size = 24),
    #legend.position = c(0.85, 0.83),  # adjust x, y for placement
    legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 18, face = "bold"),
    legend.text = element_text(size = 17),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 20, color = "black"),
    axis.text.y = element_text(size = 24, color = "black"),
    axis.title.y = element_text(size = 26, face = "bold")
); prot_plot

ggsave("../output/02-Physiology/Protein/Protein_KBay.png", 
       plot = prot_plot, width = 20, height = 8, units = "in", dpi = 300)
ggsave("../output/02-Physiology/Protein/Protein_KBay.pdf", 
       plot = prot_plot, width = 20, height = 8, units = "in", device = "pdf", dpi = 300)
```



