---
title: "Surface area calculations"
author: "Jill Ashey"
date: "2026-04-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggpubr) 
```

Read in colony metadata
```{r}
metadata <- read.csv("../data/02-Physiology/Pairs_metadata.csv")
metadata$ColonyID <- as.character(metadata$ColonyID)
```

## Wax dipping method 

Load wax data and calculate change in weight 
```{r}
wax_data <- read.csv("../data/02-Physiology/SurfaceArea/wax_surface_area.csv")
wax_data <- wax_data %>%
  mutate(delta.mass.g = With_wax_weight_g - Without_wax_weight_g)
```

Merge with colony metadata 
```{r}
wax_data <- left_join(wax_data, metadata, by = "ColonyID")
```

Calculate surface area for standards
```{r}
wax_data_std <- wax_data %>%
  filter(Sample == "Standard") %>%
  mutate(radius = Diameter/2,
         surface.area.cm2 = (pi * Diameter * Height) + (pi * radius^2))

ggplot(wax_data_std, aes(x = delta.mass.g, y = surface.area.cm2)) +
  geom_point(size = 2) +
  # Add the linear regression line for each facet
  geom_smooth(method = "lm", formula = y ~ x, color = "blue", se = FALSE) +
  # Create a separate plot for each Timepoint
  facet_wrap(~Timepoint, scales = "free") + 
  theme_bw() +
  labs(
    title = "Wax Standard Curves by Timepoint",
    x = "Wax Weight Gain (g)",
    y = "Calculated Surface Area (cm2)"
  )
```

Calculate slopes (forcing the origin) and join with sample data
```{r}
wax_data_samp <- wax_data_std %>%
  group_by(Timepoint) %>%
  nest() %>%
  mutate(model = map(data, ~lm(surface.area.cm2 ~ delta.mass.g + 0, data = .x))) %>%
  right_join(wax_data %>% filter(Sample == "Coral"), by = "Timepoint") %>%
  group_by(Timepoint) %>%
  mutate(surface.area.cm2 = predict(model[[1]], newdata = cur_data())) %>%
  select(-data, -model) 
```

Keep only colony and surface area info
```{r}
wax_data_samp <- wax_data_samp %>%
  dplyr::select(ColonyID, Species, Phenotype, Timepoint, Without_wax_weight_g, With_wax_weight_g, delta.mass.g, surface.area.cm2, Notes)
```

Save as csv 
```{r}
write.csv(wax_data_samp, "../output/02-Physiology/SurfaceArea/SurfaceArea_wax_calculations.csv")
```

## Comparison of wax dipping and fraggrameter surface areas

Read in wax calculations
```{r}
wax <- read.csv("../output/02-Physiology/SurfaceArea/SurfaceArea_wax_calculations.csv")
wax$ColonyID <- as.character(wax$ColonyID)
wax <- wax[,-1] # remove extra weird column 

# Select specific cols 
wax <- wax %>%
  select(ColonyID, Species, Phenotype, Timepoint, surface.area.cm2, Notes)
```

Read in fraggrameter SA data
```{r}
frag <- read.csv("../data/02-Physiology/SurfaceArea/fraggrameter_surface_area.csv")
frag$ColonyID <- as.character(frag$ColonyID)

# Assign dates to timepoint and select specific cols
frag <- frag %>%
  mutate(Timepoint = case_when(
    Date == "2023-04-15" ~ "April_2023",
    Date == "2023-09-14" ~ "September_2023",
    Date == "2024-01-20" ~ "January_2024",
    Date == "2024-01-21" ~ "January_2024",
    Date == "2024-05-09" ~ "May_2024",
    Date == "2024-05-10" ~ "May_2024",
    Date == "2024-10-01" ~ "October_2024",
    Date == "2024-10-02" ~ "October_2024",
    Date == "2025-01-09" ~ "January_2025",
    Date == "2025-05-08" ~ "May_2025",
    Date == "2023-09-15" ~ "September_2023",
    Date == "2025-10-03" ~ "October_2025",
    Date == "2026-01-25" ~ "January_2025")) %>%
  select(ColonyID, surface_area_cm2_fraggrameter, Timepoint, Notes)

# Merge with meta
#frag <- left_join(frag, metadata, by = "ColonyID")
```

Merge wax and frag data by timepoint and colonyID
```{r}
both <- left_join(wax, frag, by = c("ColonyID", "Timepoint"))
```

Compare 
```{r}
ggplot(both, aes(x = surface.area.cm2, y = surface_area_cm2_fraggrameter)) +
  geom_point(aes(color = Phenotype), alpha = 0.7, size = 2.5) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray50") +
  stat_regline_equation(
    aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~")),
    label.x = -Inf,       # Places text on the absolute left edge
    label.y = Inf,        # Places text on the absolute top edge
    hjust = -0.1,         # Slightly nudges text inwards away from the left border
    vjust = 1.5           # Slightly nudges text downwards away from the top border
  ) +
  facet_grid(Species ~ Timepoint) +
  labs(
    title = "Surface Area Measurement Comparison",
    x = "Wax Dipped Surface Area (cm²)",
    y = "Fraggrameter Surface Area (cm²)",
    color = "Phenotype"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(face = "bold", size = 11),
    legend.position = "bottom"
  )
```

Save as csv 
```{r}
write.csv(both, "../output/02-Physiology/SurfaceArea/SurfaceArea_both_methods.csv")
```




